Effect of quality-adjusted production function inputs

### 2015-05-08 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
Added residual plots for all models.

<<Data, eval=TRUE>>=
dataPT <- filter(IST, Country == "PT")
dataUK <- filter(Leeds, Country == "UK")
historicalData <- plyr::rbind.fill(dataPT, dataUK)
@

<<Time Series Variables>>=
# This block of code sets up the time-series variables 
# to be used for each country.
# Adding this level of abstraction makes it easier 
# for us to change variables later.
# It also makes it easier to see what is going on 
# in each formula throughout the document.
# Use of substitute establishes an R "name"
# that can be a component of a formula.
# Economic output
y_PT <- substitute(iGDP)
y_UK <- substitute(iGDP)
# Portugal undajusted inputs
k_PT_unadj <- substitute(iKstkS.L)
l_PT_unadj <- substitute(iL)
e_PT_unadj <- substitute(iXpMP)
# Portugal quality-adjusted inputs
k_PT_adj <- substitute(iKservS.L)
l_PT_adj <- substitute(ihLest)
e_PT_adj <- substitute(iUMP)
# UK unadjusted inputs
k_UK_unadj <- substitute(iKstkO.WwithRD)
l_UK_unadj <- substitute(iL)
e_UK_unadj <- substitute(iXp)
# UK quality-adjusted inputs
k_UK_adj <- substitute(iKservO.WwithRD)
l_UK_adj <- substitute(ihLest)
e_UK_adj <- substitute(iU)
# Time variables
t_PT <- substitute(iYear)
t_UK <- substitute(iYear)
@

@@

@@

<<Fitted_Data>>=
  extract_fitted_comparison <- function(mod, id){
    df <- mod$data[ , c("Year", "iGDP", "Country")]
    # Trim fitted to be same length as df to eliminate problems with possible NAs at end of fitted
    df$fitted <- yhat(mod)[c(1:length(df$Year))]
    # Parse the id to find out whether this is quality-adjusted data
    id_list <- strsplit(id, split = "[.]")[[1]]
    if (id_list[[2]] == "unadj"){
      df$adjusted <- "Unadjusted"
    } else if (id_list[[2]] == "adj"){
      df$adjusted <- "Quality-adjusted"
    } else {
      stop("Unknown adjustment.")
    }
    if (id_list[[3]] == "noE"){
      df$energy <- "Without energy"
    } else if (id_list[[3]] == "withE"){
      df$energy <- "With energy"
    } else {
      stop("Unknown energy.")
    }
    df$model <- id_list[[4]]    
    return(df)
  }
  fitted_data <- leaf_apply(models, extract_fitted_comparison, class=c("CDEmodel", "cesEst", "LINEXmodel"))
  fitted_data <- do.call(rbind, fitted_data)
@

@@

@@

@@

@@

@@

@@

@@

>>>>

<<Data, eval=TRUE>>=
dataPT <- filter(IST, Country == "PT")
dataUK <- filter(Leeds, Country == "UK")
@

<<Time Series Variables>>=
# This block of code sets up the time-series variables 
# to be used for each country.
# Adding this level of abstraction makes it easier 
# for us to change variables later.
# It also makes it easier to see what is going on 
# in each formula throughout the document.
# Use of substitute establishes an R "name"
# that can be a component of a formula.
# Economic output
y_PT <- substitute(iGDP)
y_UK <- substitute(iGDP)
# Portugal undajusted inputs
k_PT_unadj <- substitute(iKstkS.L)
l_PT_unadj <- substitute(iL)
e_PT_unadj <- substitute(iXpMP)
# Portugal quality-adjusted inputs
k_PT_adj <- substitute(iKservS.L)
l_PT_adj <- substitute(ihLest)
e_PT_adj <- substitute(iUMP)
# UK unadjusted inputs
k_UK_unadj <- substitute(iKstkO.WwithRD)
l_UK_unadj <- substitute(iL)
e_UK_unadj <- substitute(iXp)
# UK quality-adjusted inputs
k_UK_adj <- substitute(iKservO.WwithRD)
l_UK_adj <- substitute(ihLest)
e_UK_adj <- substitute(iU)
# Time variables
t_PT <- substitute(iYear)
t_UK <- substitute(iYear)

# Create some pre-econometric ratios
# Unadjusted variables
dataPT <- transform(dataPT, kPerGDP_unadj = eval(k_PT_unadj)/iGDP)
dataPT <- transform(dataPT, lPerGDP_unadj = eval(l_PT_unadj)/iGDP)
dataPT <- transform(dataPT, ePerGDP_unadj = eval(e_PT_unadj)/iGDP)
dataUK <- transform(dataUK, kPerGDP_unadj = eval(k_UK_unadj)/iGDP)
dataUK <- transform(dataUK, lPerGDP_unadj = eval(l_UK_unadj)/iGDP)
dataUK <- transform(dataUK, ePerGDP_unadj = eval(e_UK_unadj)/iGDP)
# Quality-adjusted variables
dataPT <- transform(dataPT, kPerGDP_adj = eval(k_PT_adj)/iGDP)
dataPT <- transform(dataPT, lPerGDP_adj = eval(l_PT_adj)/iGDP)
dataPT <- transform(dataPT, ePerGDP_adj = eval(e_PT_adj)/iGDP)
dataUK <- transform(dataUK, kPerGDP_adj = eval(k_UK_adj)/iGDP)
dataUK <- transform(dataUK, lPerGDP_adj = eval(l_UK_adj)/iGDP)
dataUK <- transform(dataUK, ePerGDP_adj = eval(e_UK_adj)/iGDP)

# Now create the historicalData data frame
historicalData <- plyr::rbind.fill(dataPT, dataUK)
@

@@

<<Pre-econometric graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Pre-econometric data.", warning=FALSE>>=
# Melt the data
mdata_preeconometric <- melt(historicalData, 
                             measure.vars = c("kPerGDP_unadj", "lPerGDP_unadj", "ePerGDP_unadj",
                                              "kPerGDP_adj", "lPerGDP_adj", "ePerGDP_adj"),
                             variable.name = "variable",
                             value.name = "value")
# Add a column to say whether the variable is Unadjusted or Quality-adjusted based on variable name
mdata_preeconometric$adjusted <- lapply(mdata_preeconometric$variable, function(x) {
  suffix <- unlist(strsplit(as.character(x), split="[_]")) [2]
  if (suffix == "unadj"){
    return("Unadjusted")
  } else if (suffix == "adj"){
    return("Quality-adjusted")
  } else {
    stop(paste("Unknown suffix:", suffix))
  }
})
# Fix levels
mdata_preeconometric$adjusted <- relevelFactor(as.factor(as.character(mdata_preeconometric$adjusted)), 
                                               c("Unadjusted", "Quality-adjusted"))
mdata_preeconometric$Country <- relevelFactor(as.factor(mdata_preeconometric$Country), 
                                              c("PT", "UK"))
# Now make the plot
historicalPlot(mdata_preeconometric, 
               mapping=aes(x=Year, y=value, group=variable, linetype=variable), 
               facet_formula = Country ~ adjusted, 
               line_types = c(2,4,3, 2,4,3)) + 
  xlab("") + 
  ylab("Indexed (1960=1)")
@

@@

<<Fitted_and_Residuals>>=
extract_fitted_and_resid <- function(mod, id){
  df <- mod$data[ , c("Year", "iGDP", "Country")]
  # Trim fitted to be same length as df to eliminate problems with possible NAs at end of fitted
  df$fitted <- yhat(mod)[c(1:length(df$Year))]
  df$Residuals <- resid(mod)[c(1:length(df$Year))]
  # Parse the id to find out whether this is quality-adjusted data
  id_list <- strsplit(id, split = "[.]")[[1]]
  if (id_list[[2]] == "unadj"){
    df$adjusted <- "Unadjusted"
  } else if (id_list[[2]] == "adj"){
    df$adjusted <- "Quality-adjusted"
  } else {
    stop("Unknown adjustment.")
  }
  if (id_list[[3]] == "noE"){
    df$energy <- "Without energy"
  } else if (id_list[[3]] == "withE"){
    df$energy <- "With energy"
  } else {
    stop("Unknown energy.")
  }
  df$model <- id_list[[4]]    
  return(df)
}
fitted_and_resid_data <- leaf_apply(models, extract_fitted_and_resid, class=c("CDEmodel", "cesEst", "LINEXmodel"))
# Create a data frame.
fitted_and_resid_data <- do.call(rbind, fitted_and_resid_data)

# Relevel factors before plotting
fitted_and_resid_data$Country <- relevelFactor(as.factor(fitted_and_resid_data$Country), c("PT", "UK"))
fitted_and_resid_data$adjusted <- relevelFactor(as.factor(fitted_and_resid_data$adjusted), c("Unadjusted", "Quality-adjusted"))
fitted_and_resid_data$energy <- relevelFactor(as.factor(fitted_and_resid_data$energy), c("Without energy", "With Energy"))
@

@@

%++++++++++++++++++++++++++++++
\subsection{Fits to historical data} 
\label{sec:fits_to_historical_data}
%++++++++++++++++++++++++++++++

This section shows graphs of fits to historical data and residuals.

<<Fitted and Residual Graphs>>=
fittedGraph <- function(data){
  ggplot() +
    geom_line(data=filter(data, energy=="Without energy"), mapping=aes(x=Year, y=iGDP)) +
    geom_line(data=data, mapping=aes(x=Year, y=fitted, linetype=energy)) +
    scale_linetype_manual(values=c(2,3), name="") +
    facet_grid(Country ~ adjusted) +
    xy_theme()
}

residualPlot <- function(data){
  ggplot() +
    geom_line(data=data, mapping=aes(x=Year, y=Residuals, linetype=energy)) +
    scale_linetype_manual(values=c(2,3), name="") +
    facet_grid(Country ~ adjusted) +
    xy_theme()
}
@

<<Graphs CD-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Cobb-Douglas models. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CD"))
@

<<Graphs CD-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Residuals for Cobb-Douglas models.">>=
residualPlot(filter(fitted_and_resid_data, model=="CD"))
@

<<Graphs CESkle-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($kl$)$e$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CESkle" | model=="CESkl"))
@

<<Graphs CESkle-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Residuals for CES models with $(kl$)$e$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CESkle" | model=="CESkl"))
@

<<Graphs CESlek-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($le$)$k$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CESlek" | model=="CESkl"))
@

<<Graphs CESlek-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Residuals for CES models with $(le$)$k$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CESlek" | model=="CESkl"))
@

<<Graphs CESekl-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($ek$)$l$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CESekl" | model=="CESkl"))
@

<<Graphs CESekl-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Residuals for CES models with $(ek$)$l$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CESekl" | model=="CESkl"))
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Analysis}
\label{sec:analysis}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Table~\ref{tab:sse_CES-kle_PT} shows the $sse$ values for each of these models.

\begin{table} \caption{PT $sse$ with CES and ($kl$)$e$ nesting.} 
\label{tab:sse_CES-kle_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$PT$unadj$noE$CESkl)$sse} & \Sexpr{EconModels:::sse(models$PT$adj$noE$CESkl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$PT$unadj$withE$CESkle)$sse} 
                                                            & \Sexpr{EconModels:::sse(models$PT$adj$withE$CESkle)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

Table~\ref{tab:sse_CES-lek_PT} shows the $sse$ values for each of these models.

\begin{table} \caption{PT $sse$ with CES and ($le$)$k$ nesting.} 
\label{tab:sse_CES-lek_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$PT$unadj$noE$CESkl)$sse} & \Sexpr{EconModels:::sse(models$PT$adj$noE$CESkl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$PT$unadj$withE$CESlek)$sse} 
                                                            & \Sexpr{EconModels:::sse(models$PT$adj$withE$CESlek)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

Table~\ref{tab:sse_CES-ekl_PT} shows the $sse$ values for each of these models.

\begin{table} \caption{PT $sse$ with CES and ($ek$)$l$ nesting.} 
\label{tab:sse_CES-ekl_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$PT$unadj$noE$CESkl)$sse} & \Sexpr{EconModels:::sse(models$PT$adj$noE$CESkl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$PT$unadj$withE$CESekl)$sse} 
                                                            & \Sexpr{EconModels:::sse(models$PT$adj$withE$CESekl)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

Table~\ref{tab:sse_CES-kle_UK} shows the $sse$ values for each of these models.

\begin{table} \caption{UK $sse$ with CES and ($kl$)$e$ nesting.} 
\label{tab:sse_CES-kle_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$UK$unadj$noE$CESkl)$sse} & \Sexpr{EconModels:::sse(models$UK$adj$noE$CESkl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$UK$unadj$withE$CESkle)$sse} 
                                                            & \Sexpr{EconModels:::sse(models$UK$adj$withE$CESkle)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

Table~\ref{tab:sse_CES-ekl_UK} shows the $sse$ values for each of these models.

\begin{table} \caption{UK $sse$ with CES and ($le$)$k$ nesting.} 
\label{tab:sse_CES-lek_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$UK$unadj$noE$CESkl)$sse} & \Sexpr{EconModels:::sse(models$UK$adj$noE$CESkl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$UK$unadj$withE$CESlek)$sse} 
                                                            & \Sexpr{EconModels:::sse(models$UK$adj$withE$CESlek)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

Table~\ref{tab:sse_CES-ekl_UK} shows the $sse$ values for each of these models.

\begin{table} \caption{UK $sse$ with CES and ($ek$)$l$ nesting.} 
\label{tab:sse_CES-ekl_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$UK$unadj$noE$CESkl)$sse} & \Sexpr{EconModels:::sse(models$UK$adj$noE$CESkl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$UK$unadj$withE$CESekl)$sse} 
                                                            & \Sexpr{EconModels:::sse(models$UK$adj$withE$CESekl)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}
